Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-09 examples.tex
convert examples to new ghost map token syntax

%\begin{comment}
%\todo[inline]{Identity mappings are difficult, and our current approach won't quite work. Consider trying to have a virtual pointsto for an actual page table entry (i.e., that one could use to update a page table mapping), while also having a virtual pointsto for an address that entry mapped. With the current (let's call it v1) solution, we can't actually have both of those simultaneously!  That's because the PTE pointsto will assert full ownership of the physical memory cell holding the PTE as its data value, while the virtual pointsto for the data mapped by that entry will \emph{also} assert (fractional) ownership of all entries a page table walk would traverse.
%}
%\todo[inline,color=violet]{This doesn't seem to cause issues with the mapping/unmapping examples, only with changing intermediate page table pointers. The mapping example requires a virtual pointsto for the blank PTE, and once filled in that ownership can be immediately split to create the 512 new virtual pointsto assertions for the newly mapped page. Conversely, for unmapping we'd assume ownership of all the relevant virtual pointsto assertions for the page we're unmapping, at which point we can (with a bit of work) show that they all correspond to the same L1 PTE, and extract the 512 fractional shares of that entry from the pointsto assertions.  But changing intermediate page tables, as one would do for coallescing or splitting a superpage while preserving the virtual-to-physical mappings, couldn't be done without some really complicated separating implication tricks.}
%\todo[inline,color=green]{One possible approach to resolving this, which we came up with in our Tuesday meeting, is to recognize that the current (v1) virtual points-to is too strong, because it really doesn't care about \emph{owning} those fractional resources, it only cares that \emph{something} ensures the correct page table walk exists. Iris has a ghost map resource where authoritative ownership of an individual key-value pair can be handled as a resource.  (Colin was using this in the filesystem cache.)
%We can use that mechanism to separate the virtual-to-physical translation from the physical memory involved (Kolanski and Klein may have done something similar for different reasons): (fractional) virtual points-to assertions can be defined in terms of (fractional) ownership of these authoritative ghost map entry assertions, plus sharing an invariant that the current installed page table respects all entries of the mapping. Unmapping collects the authoritative map kvpairs from collecting the assertions, and then can remove them from the ghost map and update the page tables. Critically, physical ownership of the page tables then lives in the invariant on the current page table, so some virtual pointsto assertions can refer to memory in those page tables.
%This still works with the modality, since that invariant is also semantically a predicate on a page table root.
%Let's call this v2.
%}
%\end{comment}
\subsection{Unmapping a Page}
The reverse operation, unmapping a designated page that is currently mapped,
would essentially be the reverse of
the reasoning around line 22 above: given the virtual points-to assertions for all 512
machine words of memory that the L1 entry would map,
and information about the physical location, 
full permission on the L1 entry could be obtained, allowing the construction of a
full virtual PTE pointer for it, setting to 0, and reclaiming the now-unmapped physical memory.

@@

$\specline{\mathsf{entry} \in \Xi \ast \sumapaces\rtv\delta  }_{\rtv}$

@@

$\specline{ \ulcorner \Xi \;!!\; \mathsf{entry} = \mathsf{Some V} \urcorner \ast \sumapaces\rtv\delta  }_{\rtv}$

@@

$\specline{ \sumapaces\rtv\delta  }_{\rtv}$

@@

$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \} }_{\rtv}$

@@

$\specline{\sumapaces\rtv\delta \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$

@@

\caption{Making sure that \textsf{entry}'s frame is referencing to an address backed up by a physical page.}

@@

As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the virtual memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 13 in Figure \ref{fig:calltopteinitialize}). We need to construct this virtual mapping for the physical address \textsf{entry}.
\subsubsection{Identity Mappings}
\label{subsec:identitymappings}
Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
with DMA devices.}
\todo{ismail: CHnage the text!  it is not as simple as relating paddrs existence type, more concrete relation on pfn and fresly allocated addr }
\begin{figure*}
\footnotesize
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast
   \exists \textsf{freshaddr, entry} \textsf{entry\_pfn} entry freshaddr = paddr \ast freshaddr \mapsto_{\textsf{p}} \textsf{w0\_64}  \leftrightarrow \mathsf{val} = \mathsf{w0\_64}\urcorner \; \ast\\
   \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad   \underbrace{\sumwalkabs\vaddr\qfrac\paddr }_\text{Ghost translation} \ast 
% \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
  \textsf{ where } \\
   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
   \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}

@@

Concretely speaking, when we focus on the specfication of Line 13 in Figure \ref{fig:calltopteinitialize} in Figure \ref{fig:lookingupidpampping}, to read the value referenced by physical address \textsf{entry}, as shown in the first specification Line in Figure \ref{fig:lookingupidpampping}, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the fractional ownership of the ghost translation and physical pointsto
\[\sumwalkabsent \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
via using the capability $\Xi \;!!\; \textsf{entry} = \textsf{Some V}$ and left the invariant as in the state $\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi\setminus\{\mathsf{entry}\},m)$ , and construct the virtual mapping with the resources peeled-out of the identity mapping $\Xi$ 
\[ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \]
under the side condition (*)
\begin{equation*}
 \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
\end{equation*}
assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated ad Line 19in Figure \ref{fig:calltopteinitialize}.
 \begin{figure*}
   \footnotesize
   \[
 \begin{array}{l}
   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \ulcorner \Xi !! \textsf{entry} = \textsf{Some V} \urcorner  }_{\rtv}  \\
\sqsubseteq \\
   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}\\
   \specline{ \sumapaces\rtv\delta  }_{\rtv} \\
\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
\mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
 \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} }_{\rtv}\\
   \specline{ \sumapaces\rtv\delta  }_{\rtv} \\
\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
 \end{array}
 \]
 \vspace{-1em}
\caption{Peel-out the resources needed for the virtual-pte pointsto relation for physical address \textsf{entry} to be accessed in Line 7 in Figure \ref{fig:calltopteinitialize}. }
  \label{fig:lookingupidpampping}
 \end{figure*}

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\sumapaces\rtv\delta }_{\rtv}$
;;void pte_initialize(pte_t **table) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64}\ast\sumapaces\rtv\delta }_{\rtv}$
mov    rax,-0x10[rbp] 
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \sumapaces\rtv\delta }_{\rtv}$  
;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \sumapaces\rtv\delta }_{\rtv}$  
;;return;
;;..
;;clean up the stack
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
\todo{ISMAIL: Change the following paragraphs and LInes 33 35  in \ref{fig:calltopteinitialize}, because wpdp2V is where we only know the relation between entry->pfn nextpaddr , consequently zeroed. Not here }
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 33 and 35 in Figure \ref{fig:calltopteinitialize}. There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}. First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource. Second, we see the identity mapping for \textsf{nextpaddr}, first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} . The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.

@@

%$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$ 

@@

Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization
map ($\theta$) to change its state from unmapped ($\ulcorner \theta
\;!!\;\vaddr = \texttt{None}\urcorner$ (Line 28) and the
precondition) to mapped (Line 29) using Iris' ghost-map update
($\sqsubseteq$), and construct a PTE -points-to $\vaddr
\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Line 31) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
page-table-walk insertions, $\sumapaces\rtv\delta$ from unfolding the
definition of $\mapsto_{\textsf{vpte}}$, and $\fpaddr
\mapsto_{\textsf{p}} \textsf{wzero 64}$.
\end{comment}
Finally, by existentially quantifying away \textsf{fpaddr},
we can shift the PTE points-to into a normal virtual points-to.

>>>>

\todo[inline]{need new text explaining both mapping and the page table walk}

%\begin{comment}
%\todo[inline]{Identity mappings are difficult, and our current approach won't quite work. Consider trying to have a virtual pointsto for an actual page table entry (i.e., that one could use to update a page table mapping), while also having a virtual pointsto for an address that entry mapped. With the current (let's call it v1) solution, we can't actually have both of those simultaneously!  That's because the PTE pointsto will assert full ownership of the physical memory cell holding the PTE as its data value, while the virtual pointsto for the data mapped by that entry will \emph{also} assert (fractional) ownership of all entries a page table walk would traverse.
%}
%\todo[inline,color=violet]{This doesn't seem to cause issues with the mapping/unmapping examples, only with changing intermediate page table pointers. The mapping example requires a virtual pointsto for the blank PTE, and once filled in that ownership can be immediately split to create the 512 new virtual pointsto assertions for the newly mapped page. Conversely, for unmapping we'd assume ownership of all the relevant virtual pointsto assertions for the page we're unmapping, at which point we can (with a bit of work) show that they all correspond to the same L1 PTE, and extract the 512 fractional shares of that entry from the pointsto assertions.  But changing intermediate page tables, as one would do for coallescing or splitting a superpage while preserving the virtual-to-physical mappings, couldn't be done without some really complicated separating implication tricks.}
%\todo[inline,color=green]{One possible approach to resolving this, which we came up with in our Tuesday meeting, is to recognize that the current (v1) virtual points-to is too strong, because it really doesn't care about \emph{owning} those fractional resources, it only cares that \emph{something} ensures the correct page table walk exists. Iris has a ghost map resource where authoritative ownership of an individual key-value pair can be handled as a resource.  (Colin was using this in the filesystem cache.)
%We can use that mechanism to separate the virtual-to-physical translation from the physical memory involved (Kolanski and Klein may have done something similar for different reasons): (fractional) virtual points-to assertions can be defined in terms of (fractional) ownership of these authoritative ghost map entry assertions, plus sharing an invariant that the current installed page table respects all entries of the mapping. Unmapping collects the authoritative map kvpairs from collecting the assertions, and then can remove them from the ghost map and update the page tables. Critically, physical ownership of the page tables then lives in the invariant on the current page table, so some virtual pointsto assertions can refer to memory in those page tables.
%This still works with the modality, since that invariant is also semantically a predicate on a page table root.
%Let's call this v2.
%}
%\end{comment}
\subsection{Unmapping a Page}
\todo[inline]{this must move, must come after mapping}
The reverse operation, unmapping a designated page that is currently mapped,
would essentially be the reverse of
the reasoning around line 22 above: given the virtual points-to assertions for all 512
machine words of memory that the L1 entry would map,
and information about the physical location, 
full permission on the L1 entry could be obtained, allowing the construction of a
full virtual PTE pointer for it, setting to 0, and reclaiming the now-unmapped physical memory.

@@

$\specline{\mathsf{entry} \in \Xi \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

@@

$\specline{ \ulcorner \Xi \;!!\; \mathsf{entry} = \mathsf{Some V} \urcorner \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

@@

$\specline{ \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$

@@

$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \} }_{\rtv}$

@@

$\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \mathsf{nextpaddr} \in \Xi\setminus \{ \mathsf{entry} \} \cup \{ \mathsf{nextpaddr} \} }_{\rtv}$

@@

\caption{Ensuring \textsf{entry} points to a valid next table, allocating if necessary.}

@@

As we see in Figure \ref{fig:calltopteinitialize}, we have the physical address (\textsf{entry}) of the page table entry kept in the register $\texttt{r8}$, but not the virtual memory mapping to load the entry value referenced by the physical address \textsf{entry} (Line 13 in Figure \ref{fig:calltopteinitialize}). We need to construct this virtual mapping for the physical address \textsf{entry}.
\subsubsection{Identity Mappings}
\label{subsec:identitymappings}
Most kernels maintain an invariant on their page tables that the virtual address of any page used for a page table lives at a virtual address
whose value is a constant offset from the physical address a practice sometimes referred to as \emph{identity mapping} (even though the physical-to-virtual translation
is typically not literally the identity function, but adding a constant offset).\footnote{Some kernels do this for all physical memory on the machine, simplifying interaction
with DMA devices.}
\todo{ismail: CHnage the text!  it is not as simple as relating paddrs existence type, more concrete relation on pfn and fresly allocated addr }
\begin{figure*}
\footnotesize
\todo[inline]{Ismail, look at this figure, something went weird (was already like this, I didn't add this)}
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)\stackrel{\triangle}{=}\\ \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \ast 
  \bigast{(\vaddr, \textsf{paddr})\in \ptablestore}{\exists\;(\textsf{l4e, l3e, l2e, l1e, paddr})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr\textsf{, l4e, l3e, l2e, l1e, paddr})} \\
  \bigast{(\paddr,\mathsf{T}) \in \Xi}{\exists\; (\textsf{qfrac, q, val,}\vaddr) \ldotp \ulcorner \vaddr = \paddr + \textsf{KERNBASE} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast
   \exists \textsf{freshaddr, entry} \textsf{entry\_pfn} entry freshaddr = paddr \ast freshaddr \mapsto_{\textsf{p}} \textsf{w0\_64}  \leftrightarrow \mathsf{val} = \mathsf{w0\_64}\urcorner \; \ast\\
   \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad   \underbrace{\fracghostmaptoken{\delta}{\vaddr}{\paddr}{\qfrac} }_\text{Ghost translation} \ast 
% \qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad    
\underbrace{\paddr \mapsto_{\mathsf{p}}\{\textsf{qfrac}\}\; \vale}_\text{Physical location}}\\
  \textsf{ where } \\
   \textsf{ASpace\_Lookup}_{\textsf{id}}(\ptablestore,\Xi,m) \stackrel{\triangle}{=} \lambda\textsf{ cr3val} \ldotp \; \exists \gammaPred \; \ldotp \ulcorner m \; !!\; \textsf{cr3val} = \textsf{Some } \gammaPred \urcorner \ast
   % \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast  \ownGhostpv\gammaPred{\authfull{\pvmapping\Xi}}
   \ptableabswalk{\delta,\ptablestore} \ast \pvmapping{\delta,\Xi}

@@

Concretely speaking, when we focus on the specfication of Line 13 in Figure \ref{fig:calltopteinitialize} in Figure \ref{fig:lookingupidpampping}, to read the value referenced by physical address \textsf{entry}, as shown in the first specification Line in Figure \ref{fig:lookingupidpampping}, we open our the addresss-space invariant ($\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi,m)$), and then obtain the fractional ownership of the ghost translation and physical pointsto
% \[\sumwalkabsent \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
\[\fracghostmaptoken{\delta}{(\texttt{entry + KERNBASE})}{\textsf{entry}}{\textsf{qfrac}} \ast \mathsf{entry}\mapsto_{\mathsf{p,qfrac}}\; \vale \]
by using the capability $\Xi \;!!\; \textsf{entry} = \textsf{Some V}$ and left the invariant as in the state $\mathcal{I}\textsf{ASpace}_{\textsf{id}}(\ptablestore,\Xi\setminus\{\mathsf{entry}\},m)$ , and construct the virtual mapping with the resources peeled-out of the identity mapping $\Xi$ 
\[ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \]
under the side condition (*)
\begin{equation*}
 \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \tag{*}
\end{equation*}
assuring that looking the identity mapping for \textsf{entry} is safe under the subtle justification which equates the full ownership to the non/presence of the entry which can only be known when investigated ad Line 19in Figure \ref{fig:calltopteinitialize}.
 \begin{figure*}
   \footnotesize
   \[
 \begin{array}{l}
   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry} \textsf{\_} \ast \ulcorner \Xi !! \textsf{entry} = \textsf{Some V} \urcorner  }_{\rtv}  \\
\sqsubseteq \\
   \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
\mathsf{mov} \;   [ \mathsf{r8} ],\mathsf{rdi}\\
 \specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\setminus \{ \mathsf{entry} \},m)  \ast \textsf{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{entry}  \ast  \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv} \\
\specline{ \textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry entry\_val} \ast \ulcorner  \textsf{qfrac} = 1 \leftrightarrow \; \lnot(\textsf{entry\_present entry\_val})\urcorner}_{\rtv} \\
 \end{array}
 \]
 \vspace{-1em}
\caption{Peel-out the resources needed for the virtual-pte pointsto relation for physical address \textsf{entry} to be accessed in Line 7 in Figure \ref{fig:calltopteinitialize}. }
  \label{fig:lookingupidpampping}
 \end{figure*}

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
;;void pte_initialize(pte_t **table) {
...
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64}\ast\ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$
mov    rax,-0x10[rbp] 
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi \;!!\; \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$  
;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr}  \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \textsf{nextpaddr} \mapsto_{\textsf{p}} \textsf{w0\_64} \ast \ghostmaptoken{\delta{}s}{\rtv}{\delta} }_{\rtv}$  
;;return;
;;..
;;clean up the stack
\end{lstlisting}
\vspace{-1em}
\caption{Allocating a physical page }
\label{pteinitializespec}
\end{figure}
\todo{ISMAIL: Change the following paragraphs and LInes 33 35  in \ref{fig:calltopteinitialize}, because wpdp2V is where we only know the relation between entry->pfn nextpaddr , consequently zeroed. Not here }
Since we are using \textsf{pte\_initialize} for page-table address allocation, we must relate this newly allocated physical address to the identity mapping map $\Xi$ -- Specification Lines 33 and 35 in Figure \ref{fig:calltopteinitialize}. There are two important aspects on relating a freshly allocated page-table physical address to the identity-mapping which we focus in more detail separately in the specification of \textsf{pte\_initialize} in Figure \ref{fig:pteinitializespec}. First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource. Second, we see the identity mapping for \textsf{nextpaddr}, first not appearing in the identity map (Specification Line 11 in Figure \ref{fig:pteinitializespec}), then as inserted into the identity-map with a value \textsf{Z} ($\Xi \cup \{\mathsf{nextpaddr}\}$) together with the fragmental capability  -- $\ulcorner \Xi \; !!\; \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- at Lines 13 and 14 in Figure \ref{fig:pteinitializespec} . The value type \textsf{Z}, unlike the value type \textsf{V} denoting the random value in Line 8 in Figure \ref{fig:calltopteinitialize}, denotes \textsf{zeroed} value.

@@

%$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \ghostmaptoken{\delta{}s}{\rtv}{\delta}  }_{\rtv}$ 

@@

Finally, we can insert $\vaddr$ into the ghost page-table-walk summarization
map ($\theta$) to change its state from unmapped ($\ulcorner \theta
\;!!\;\vaddr = \texttt{None}\urcorner$ (Line 28) and the
precondition) to mapped (Line 29) using Iris' ghost-map update
($\sqsubseteq$), and construct a PTE -points-to $\vaddr
\mapsto_{\textsf{vpte}}\; \qfrac \;\fpaddr\;(\textsf{wzero 64})$ (Line 31) using $\sumwalkabs\vaddr\qfrac\fpaddr$ obtained from ghost
page-table-walk insertions, $\ghostmaptoken{\delta{}s}{\rtv}{\delta}$ from unfolding the
definition of $\mapsto_{\textsf{vpte}}$, and $\fpaddr
\mapsto_{\textsf{p}} \textsf{wzero 64}$.
\end{comment}
Finally, by existentially quantifying away \textsf{fpaddr},
we can shift the PTE points-to into a normal virtual points-to.
